\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-15 trunk/arch.tex
- Further refinements on synchronization presentation by statin many of the assumption we do in our reactor to help for quick decision - Reorganized/rewrote the end of synchronization in order to smoothly transition to deliberation/planning section

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead $\pi$  and expected
  latency $\lambda$ 
  % \comment{$\pi \&  \lambda$ both defined for the first time} 
  before
  being able to produce plan to the agent depending on where it is
  located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal.Even
in this condition, \rx requires each reactor to produce at every tick
its {\em internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

@@

We introduce this requirement as a new type of flaw for the \eu
framework defined as follow: 
\begin{definition}
  \label{def:csf}
  A {\em current state flaw} indicate that one of the reactor {\em
    internal} timelines does not provide a fully defined state value
  for the execution frontier $\tau$. Such flaw can be due to one 
  of the 2 following reasons:
  \begin{itemize}
  \item The absence of a token in this timeline which overlaps the
    instant $\tau$. 
  \item An open decision between terminating the last past token for
    this timeline or pushing its end in the future.
  \end{itemize}
\end{definition}

The core idea bewteen this flaw is to ensure that at any single tick
the {\em synchronization} solver will identifies all the {\em
  internal} timelines for which the current state value is yet to be
identified. We handle their resolution based on following
proposition. 
\begin{proposition}
  \label{prop:csf:resolve}
  A {\em current state flaw} can  be resolved using one of the
  following choices which are eveluated accroding to this order:
\begin{enumerate}

@@

\end{enumerate}
\end{proposition}
Intuitively the first choice enforces an intertail value asumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice 
favorise the adavnce of the reactor current plan if any; this direct
the {\em synchronization} decision toward assuming that its current
plan is properly executed. Finally if none of these two choices did
provide a valid solution -- meaning that both lead to an inconsittant
plan -- the reactor tries all the possible state values remaining for
this timeline until it finds one that allows it to identify its
state as required by \rx architecture.

@@

\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

The steps 2 \& 3 are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after 
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to make the
assumption that the current state value of an
{\em internal} timeline does not depend on the future or more accurately
that any choices made during this synchronization will not lead to a
future inconsistency. Such an assumption impacts the set of possible 
domains the reactor can support while remaining complete. Take for 
example this rule from our \texttt{Shopping} model :

@@

\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants to be at. This first level of
relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.  % The problem can even be relaxed further if needed but
% this relaxation gives already more felxibility for the reactor to
% recover from local failures.
As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions. Since such a process
is required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be carefully designed.

At the engine level careful choices were made in both the focus of the
search -- limited only to tokens that can potentially overlap the
current tick -- and heuristics where the internal state flaw is
considered last during the search process. The later limits the amount
of backtracking for a solution -- or at least how deep one needs to
backtrack in a decision tree. Similarly the sequence in which we
evaluate the possible resolution for such a flaw -- presented
previously -- was selected according to some simple assumptions that % takes
% benefit of a possibly assume the existing plan by assuming that 
if the outcome of synchronization appears consistent with our current
plan then it probably reflects that the plan has executed
\comment{that the past is monotonic?}. Such an assumption helps direct
the internal state evaluation and in nominal cases, proves to be a
sound solution.

One final aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

@@

The number of flaws between two steps \comment{between deliberation
  and synch.?} can evolve for the following reasons that are exogenous
to deliberation:

@@

Such a scope allows filtering of tokens that are either necessarily
ending before $\tau$ or necessarily starting after, focusing the
planning problem on only those tokens that are within a ``reasonable''
future. This window is taken into account by the \rx agent which will
notify a reactor of new {\em internal} goals not only when they
overlap this reactor specific window but also for the deliberation
solver in order to reduce the number of tokens to be evaluated to
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

@@

A core aspect of this reactor is that it takes advantage of the
specific sequencing enforced by the \rx architecture along with the
fact that only one $plan$ structure is maintained in the reactor and
shared between synchronization and planning to allow these two
processes to not only be interleaved but impact each other through the
way they manipulate this plan.

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occur in the plan. In or
presentation of synchronization, we isolated the problem by stating
that we will assume at this point that the reactorhas no future goal
or more accurateely no plan to enforce at this stage. By doing so we
were able to develop the way we augment yhe europpa solver to be able
to propagete {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. By doing so we were
able to focus on synchronization as a state evluation process that can
be resolved within the \eu framework with few extensions. simliarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning process will evolve.

We first anlyse the case during which synchronization occurs after the
planning has vcompleted (ie the last seep of deliberation resulted
into a plan that is considered as complete for the current
look-ahead). As synchronization starts with the plan produced this
will gave a frame for the search process. Therefore this deliberation
will not only be a model-based estimation uniquely based on the {\em
  external} observations but will also include the reactor intentions
as described by the plan produced. Consider for example the Shopping
Agent model described during the \eu presentation and that the
location of the \texttt{agent} is an {\em external} timeline. For
example as, the reactor add the goal to have \texttt{Own} milk and the
current {\em external} states indicated that it was \texttt{At(home)}
it produced the plan in Fig. \ref{fig:shop:exec0}. The {\em external}
part of the plan -- namely the tokens \texttt{Going(home,
  superMarket)} and \texttt{At(SuperMarket)} -- and integrated as
goals for the reactor managing this specific state varaiable {\em
  internally}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}
  \caption{Shopping example : plan produced before synchronization. 
    Tokens in red indicate observations, 
    tokens in blue the goal that this plan attempt to solve. Arrows
    indicate temporal constraints.}
  \label{fig:shop:exec0}
\end{figure}

Now, consider that at the next tick $\tau' = \tau+1$ the reactor managing the
agent location was able to change the location state varaibale to the
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation in the
exisiting plan (restricting in turn the start time of \texttt{Going}
and by propagation of the constraints \texttt{Go} to be {\em exactly}
$\tau'$). While focusing uniquely on the execution frontier ($\tau'$
in Fig. \ref{fig:shop:exec1}), the \eu solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.

>>>>

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as

  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$
  % \comment{$\pi \& \lambda$ both defined here for the first time}
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

@@

We first show how a reactor can track and identify its state in the
context where it does not have any compelling need to deliberate. We
assume that we have a reactor that has no future goal. Even in this
condition, \rx requires each reactor to produce at every tick its {\em
  internal} state. For those reactors not owning {\em internal}
timelines, \rx makes sure that {\em external} observations are taken
into account for every tick as shown in \cite{py10}. Such a
requirement is enforced to ensure that all reactors share a consistent
view of the world with advancing time.

@@

We introduce this requirement as a new type of flaw for the \eu
framework defined as follow:

\begin{definition}
  \label{def:csf}
  A {\em current state flaw} indicates that a reactor {\em internal}
  timeline does not provide a fully defined state value for the
  execution frontier $\tau$. Such a flaw can be result of either of
  the following reasons:

  \begin{itemize}
  \item The absence of a token in this timeline which overlaps the
    instant $\tau$.
  \item An open decision between terminating the last token for this
    timeline or pushing its end in the future.
  \end{itemize}

\end{definition}

The core idea between this flaw is to ensure that at any single tick
the {\em synchronization} solver will identify all the {\em internal}
timelines for which the current state value is yet to be
identified. We handle their resolution based on the following
proposition.

\begin{proposition}
  \label{prop:csf:resolve}
  A {\em current state flaw} can be resolved using one of the
  following choices which are evaluated in the following order:

@@

\end{enumerate}
\end{proposition}
Intuitively the first choice enforces an intertial value assumption
where the reactor will tend to maintain its current as long as the
model and external observations allow it. The second choice favorise
the advance of the reactor current plan if any; this direct the {\em
  synchronization} decision toward assuming that its current plan is
properly executed. Finally if none of these two choices did provide a
valid solution -- meaning that both lead to an inconsistant plan --
the reactor tries all the possible state values remaining for this
timeline until it finds one that allows it to identify its state as
required by \rx architecture.

@@

\begin{enumerate}
\item Integrate the external state as provided by the owner of each
  external timeline into the plan database.
\item Propagate this information in the plan database following the
  model $\mathcal{M}$ of this reactor.
\item Resolve the current state value of each internal timelines based
  on Proposition \ref{prop:csf:resolve}.
\end{enumerate}

%%%% Extra paragraph:  targeted to a more AI centric audience
% The step 1 translates new {\em external} observations received from
% \rx into europa tokens with a start time restricted to $\tau$ or, for
% the timelines that did not receive a new observation,  restrict the
% previous token end time to the interval $[\tau+1, \infty)$. All of
% these tokens are produced as facts and consequently cannot be 
% excluded from the plan. They are also maintained by the reactors as
% observation tokens which will allow to avoid to mistakenly remove 
% them when one of the two reactor solvers attempt to relax former
% decisions made.\comment{This relaxation is developed on the last 
%   subsection further in depth}
%%%%

Steps $2$ \& $3$ are handled by the \eu solver from algorithm
\ref{alg:europa:solve}. We only enforce that the {\em current state
  flaws} are selected last by the algorithm in order to follow the
sequencing given above. As {\em synchronization} focusses on the
execution frontier wee also filter out all flaws that are outside of
the current tick scope. This mean that if this flaw relates to a token
that is {\em necessarily} ending before $\tau$ or starting after
$\tau+1$ then the solver will ignore it. This helps to reduce the
scope of the synchronization to exclusively the current state of the
reactor and reduce dramatically the search space for the {\em
  synchronization} solver. In order to do so we need to assume 
that the current state value of an {\em internal} timeline does not 
depend on the future or more accurately that any choices made during 
this {\em synchronization} will not lead to a future inconsistency. 
Such an assumption impacts the set of possible domains the reactor 
can support while remaining complete. Take for  example this rule from 
our \texttt{Shopping} model :

@@

% \begin{minipage}[t]{\textwidth}
% \begin{figure}
\begin{verbatim}
 1 Agent::Go {
 2   met_by(condition object.location.At origin);
 3   eq(from, origin.loc);
 4
 5   contains(effect object.location.Going going);
 6   eq(going.from, from);
 7   eq(going.to, to);
 8   
 9   meets(effect object.location.At destination);
10   eq(to, destination.loc);
11   going meets destination;
12 }
\end{verbatim}
% \label{lab:model:shopping}
% \caption{\small Model fragment to enforce a \texttt{Going} end condition}
% \end{figure}
% \end{minipage}

We do so by relaxing the constraint in line $5$ from a strict
\texttt{equals} to a \texttt{contains} and to add the new constraint
at line $11$ which ensures that the last \texttt{Going} ends up in the
location the agent actually wants to be at. This first level of
relaxation will therefore allow the agent to use multiple
\texttt{Going} tokens should the first one fail to reach the
objective.  % The problem can even be relaxed further if needed but
% this relaxation gives already more flexibility for the reactor to
% recover from local failures.
As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions. Since such a process
is required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be carefully designed.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. And by doing so we allow for graceful degradation
where other reactors can readapt their plans as they receive this
observation.

Finally, at the engine level careful choices were made in 
both the focus of the search -- limited only to tokens that 
can potentially overlap the current tick -- and heuristics where 
the {\em current state flaw} is considered last during the search
process. The later limits the amount of backtracking for a solution 
-- or at least how deep one needs to backtrack in a decision tree. 
Similarly the sequence in which we evaluate the possible resolution 
for such a flaw -- presented in Proposition \ref{prop:csf:resolve} -- 
was selected according to  some simple assumptions that any external 
observation that is compatible with an existing token in the current
plan reflects the execution of the plan itself. In other words this
means that our solver always prefer to use tokens already existing in
the plan in order to identify its current internal state. By doing so
we directs the synchronization search by exploiting the structure of
the plan produced during previous deliberation steps. This implies
that a deliberation steps have been done and did impact the plan
structure and calls for a description on how our \eu reactor resolve
such steps which we are now going to describe further.

@@

The number of flaws between two steps \comment{between synch. and
  execution?} can evolve for the following reasons that are exogenous
to deliberation:

@@

\comment{Insert a figure here [AAMAS?] showing $\pi \lambda$} Such a
scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable''
future. This window is taken into account by the \rx agent which will
notify a reactor of new {\em internal} goals not only when they
overlap this reactor specific window but also for the deliberation
solver in order to reduce the number of tokens to be evaluated to
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

@@

One key aspect of the \rx framework is intertwined plan execution. As
seen above, it is synchronization which drives planning, so the
co-mingling between these two processes is critical to maintaining
state within a robot immersed in a dynamic environment. Also important
is that \rx maintains 
% A core aspect of this reactor is that it takes advantage of the
% specific sequencing enforced by the \rx architecture along with the
% fact that only 
one $plan$ structure in the reactor and shared between synchronization
and planning. This allows these two processes to not only be
interleaved but how they can impact each other via plan manipulation.

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occurs. In our presentation of 
{\em synchronization}, we isolated the problem by stating
that we will assume at this point that the reactor has no future goal
or more accurately no plan to enforce. By doing so we
were able to develop the way we augment the \eu solver to be able
to propagate {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. This allowed us 
to focus on synchronization as a state evaluation process that can
be resolved within the \eu framework with few extensions. Similarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning could evolve.

We first analyze the case during which synchronization occurs after
planning has completed (\ie the last step of deliberation results in a
plan that is considered complete for the current look-ahead). As
synchronization starts with a plan % produced this will gave a frame for
% the search process. Therefore 
this deliberation will not only be a model-derived estimation uniquely
based on {\em external} observations but will also include reactor
intentions specified in the available partial plan. Consider for
example the Shopping Agent model described where the \texttt{Agent} is
an {\em external} timeline. As the reactor adds the goal to
\texttt{Own} milk and the current {\em external} timeline indicates
that the robot was \texttt{At(home)}, it produces a plan as shown in
Fig. \ref{fig:shop:exec0}. The {\em external} part of the plan --
namely the tokens \texttt{Going(home, superMarket)} and
\texttt{At(SuperMarket)} -- are installed as goals for the reactor
managing this state variable {\em internally}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}
  \caption{\small Shopping example: A plan produced before
    synchronization.  Tokens in red indicate observations, tokens in
    blue the goal that this plan attempts to solve for. Arrows
    indicate temporal constraints.}
  \label{fig:shop:exec0}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lasting for an unknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization, the solver can identify the similarity between this
new observation and the next token in the existing plan. This implies
that instead of initiating an exhaustive search of all possible
implications, the synchronization process starts its execution with
the assumption that these new observations are consitent with the
currently maintained plan. As a consequence, at the end of
synchronization we will likely have the plan depicted in
Fig. \ref{fig:shop:exec1}. Which just propagated the information
provided by the new observation in the exisiting plan (restricting in
turn the start time of \texttt{Going} and by propagation of the
constraints \texttt{Go} to be {\em exactly} $\tau'$). While focusing
uniquely on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}), the \eu solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.
